GM: Raw Material

### 2022-11-27 supp/gmraw.qmd
<edit struc>
<edit index>

# GM - Site/Raw Material

@@

This analysis follows recent analyses of Perdiz arrow points ([Selden and Dockall, in review](https://github.com/seldenlab/perdiz3/blob/main/ms/perdiz3/perdiz3.pdf)) and Gahagan biface morphology across the southern Caddo area and central Texas [@RN11783;@RN21001;@RN11264], where Perdiz arrow points and Gahagan bifaces have been found to differ in shape across the same geography as Caddo bottles [@RN11801;@RN11782;@RN11716;@RN8312]. This analysis also follows another geometric morphometric analysis in which Perdiz arrow points from the ancestral Caddo area were found to differ as a function of _time_, _raw material_, and _burial context_ [@RN9551].

@@

@@

Principal components analysis [@RN1746] was used to visualise shape variation among the elliptical bifaces, and the scatterplot represents the dispersion of shapes in tangent space [@RN8633;@RN5616;@RN11143;@RN7550]. Shape ranges described by each principal axis are commonly visualized using thin-plate spline warping of a reference image [@RN1731;@RN479].

@@

To assess whether shape changes with size (allometry), Procrustes ANOVAs [@RN1749] were run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756]. A residual randomization permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719].

@@

To assess whether shape and size differ by `size class`, Procrustes ANOVAs [@RN1749] were run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756]. A residual randomization permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719].

@@

The trajectory analysis was followed by an analysis of morphological disparity [@RN11107;@RN7041;@RN5694] to identify potential differences in morphological diversity between size classes.

>>>>

---
title: "GM: Raw Material"
bibliography: [book.bib]
---

@@

This analysis follows recent analyses of Perdiz arrow points ([Selden and Dockall, in review](https://github.com/seldenlab/perdiz3/blob/main/ms/perdiz3/perdiz3.pdf)) and Gahagan biface morphology across the southern Caddo area and central Texas [@RN11783; @RN21001; @RN11264], where Perdiz arrow points and Gahagan bifaces have been found to differ in shape across the same geography as Caddo bottles [@RN11801; @RN11782; @RN11716; @RN8312]. This analysis also follows another geometric morphometric analysis in which Perdiz arrow points from the ancestral Caddo area were found to differ as a function of *time*, *raw material*, and *burial context* [@RN9551].

@@

# read data and define number of sLMs ----
shapes <- readShapes("shapes")
shapesGM <- readland.shapes(shapes, nCurvePts = c(15, 20, 35))

@@

Principal components analysis [@RN1746] was used to visualise shape variation among the elliptical bifaces, and the scatterplot represents the dispersion of shapes in tangent space [@RN8633; @RN5616; @RN11143; @RN7550]. Shape ranges described by each principal axis are commonly visualized using thin-plate spline warping of a reference image [@RN1731; @RN479].

@@

To assess whether shape changes with size (allometry), Procrustes ANOVAs [@RN1749] were run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756]. A residual randomization permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655; @RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719].

@@

To assess whether shape and size differ by `size class`, Procrustes ANOVAs [@RN1749] were run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756]. A residual randomization permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655; @RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719].

@@

The trajectory analysis was followed by an analysis of morphological disparity [@RN11107; @RN7041; @RN5694] to identify potential differences in morphological diversity between size classes.
